### 2024-04-19 paper/paper.qmd
completed scripts

---
title: "Physical activity is as tied to obesity as we expect it to be"
subtitle: "An analysis of American exerise habits and their BMI"
author: 
  - Victor Ma
thanks: "Code and data are available at: https://github.com/bestmustard/activity-bmi"
date: 03/16/2024
date-format: long
abstract: ""
format: pdf
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72

header-includes:
 \usepackage{float}
 \floatplacement{figure}{H}
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(dataverse)
library(knitr)
library(modelsummary)
library(rstanarm)
library(tidybayes)
library(dplyr)
library(patchwork)
library(ggplot2)
library(broom.mixed)
activity_data <- read_csv("../inputs/data/analysis_data/analysis_data.csv")
activity_data$obese = as_factor(activity_data$obese)
activity_data$activity = factor(
      activity_data$activity,
      levels = c("No Activity", "Strength", "Cardio", "Cardio + Strength", "Double Cardio")
    )
```

# Introduction

## Data Preparation and Cleaning
The 2020 CCES data was first obtained via the dataverse package directly from Harvard Dataverse, and then saved in parquet format using the arrow package for efficient storage and access [@arrow]. The cleaning process involved filtering for registered voters who chose either Biden or Trump, in order to treat presidential votes as a binary outcome.

The respondent variables (age, gender, and education) were transformed for clarity. Age was categorized into groups ("65+", "45-64", "30-44", "18-29"), gender was labeled ("Male", "Female"), and education levels were delineated ("No HS", "High school graduate", "Some college", "2-year", "4-year", "Post-grad"). These steps utilized the dplyr package for data manipulation.

The cleaned dataset was then saved in both CSV and parquet formats.

The distributions for each explanatory variable are illustrated in @fig-activity and @fig-income below:

```{r out.height="20%"}
#| label: fig-activity
#| fig-cap: "Number of respondents at each activity level"
#| echo: false
#| warning: false
#| message: false

ggplot(activity_data, aes(x = activity, fill = activity)) +
  coord_flip()+
  geom_bar() +
  scale_fill_brewer(palette = "Paired") + # More colors for distinction
  theme_minimal() +
  labs(title = "Respondents per Level of Activity",
       y = "Activity Level",
       x = "Count") +
  theme(legend.title = element_blank())

```

```{r out.height="20%"}
#| label: fig-income
#| fig-cap: "Number of respondents per income bracket"
#| echo: false
#| warning: false
#| message: false
ggplot(activity_data, aes(x = income)) +
  geom_bar(fill = "#778899") +
  theme_minimal() +
  labs(title = "Respondents per income bracket", x = "Income (Thousands)", y = "Count")+
  theme(legend.title = element_blank())

```

@fig-activity tells us that 4-year degree graduates are the most well represented in the polls, followed by high school graduates, people with some college, and then 2-year degree graduates. People who have not graduated high school are hardly represented at all.

In @fig-income we see 45-64 being the most dominant age group, with much less representation in the 18-29 range. Also, about 20% more females participated in the study than males.

Below are some figures representing who people voted for based on their age and gender, followed by education level and gender.

```{r out.height="20%"}
#| label: fig-activityobesity
#| fig-cap: "Prevalence of obesity by activity level"
#| echo: false
#| warning: false
#| message: false

# Relationship between Obesity and Activity Level
ggplot(activity_data, aes(x = activity, fill = obese)) +
  geom_bar(position = "fill") +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal() +
  labs(title = "Obesity Proportion by Activity Levels", x = "Activity Level", y = "Proportion")
```
@fig-activityobesity shows a negative correlation between votes for Joe Biden and older age groups.
```{r out.height="20%"}
#| label: fig-incomeobesity
#| fig-cap: "Prevalence of obesity by income level"
#| echo: false
#| warning: false
#| message: false
ggplot(activity_data, aes(x = income, fill = obese)) +
  geom_bar(position = "fill") +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal() +
  labs(title = "Obesity Proportion by Income Brackets", x = "Age Group", y = "Proportion")
```
We can see that in @fig-gendervotes, more women voted for Joe Biden than men.

# Model

Logistic regression is a model used when the outcome or dependent variable is binary, which fits this scenario perfectly as I am modelling the binary outcome of below 30 BMI or a BMI of 30 and above.

The regression model will calculate the log odds of the probability that a person has a BMI of 30 or above, and then map it to a probability between 0 and 1 through the logistic function.

The standard logistic function $\sigma(t)$ for a real-valued input $t$ is defined as:

```{=tex}
\[ \sigma(t) = \frac{1}{1 + e^{-t}} \]
```

The graph of the logistic function is an S-shaped curve known as a sigmoid curve. It approaches 1 as $t$ goes to positive infinity and approaches 0 as $t$ goes to negative infinity.

In logistic regression, the input $t$ is the linear combination of predictors including the intercept, which can be represented as $\beta_0 + \beta_1X_1 + \beta_2X_2 + ... + \beta_nX_n$. The logistic function then translates this into a probability that the dependent variable is 1 (has a BMI above 30).

In this situation, I will be using the predictors activity level and income and then applying the logistic function to get the probability $P(Y_i=1)$ that a respondent $i$ is considered obese.

This model is particularly strong at handling categorical dependent variables, which each of my explanatory variables fall under [@Hosmer2013].

## Model Specification

The logistic regression model is defined as:

```{=tex}
\[
\log\left(\frac{P(Y_i=1)}{1 - P(Y_i=1)}\right) = \beta_0 + \beta_1X_{\text{activity},i} + \beta_2X_{\text{income},i}
\]
```
## Model set-up

- $Y_i$ is the binary indicator of having a BMI of 30 or above (1) versus a BMI below 30 for respondent $i$.
- $X_{\text{activity},i}$, $X_{\text{income},i}$ are the activity level and income of respondent $i$, respectively.
- $\beta_0$ represents the model intercept, while $\beta_1$ and $\beta_2$ are coefficients quantifying the effects of activity level and income on the likelihood of being considered obese.

I fit my logistic regression model to the data using 'stan_glm()' function from the 'rstanarm' package in R [@citeR, @Rstanarm]. This function will automatically determine each of the $\beta$ coefficients in the model, using a smaller slice sample of 3000 from the data we processed. This function also uses Bayesian logistic regression with the default priors from 'rstanarm'.

### Model Justification

In order to better interpret the results of the model, I can create a coefficient plot to visually see the effect sizes of the predictor variables on the likelihood of an individual supporting Biden.

@fig-coefficient maps each predictor variable on the y-axis to an effect size and confidence interval on the x-axis. The effect size is the change in log-odds of supporting Biden for a one-unit increase in the predictor variable, which is essentially how much impact each demographic characterstic has an effect on voting for Biden.

The confidence intervals tell us the range within we can be confident that the true effect lies. Smaller confidence intervals means there is a higher level of precision in the estimate of the effect size. I am not as interested in the intervals that cross zero because that means that there is data to support each side (Biden or Trump) and so they are less statistically significant.

```{r}
#| label: fig-coefficient
#| fig-cap: "Coefficient plot of demographics"
#| echo: false
#| warning: false
#| message: false
obesity_likelihood <- readRDS(file ="../models/obesity_likelihood.rds")

# Assuming `tidy_obesity_likelihood` is a tidy version of your model with terms and estimate
tidy_obesity_likelihood <- tidy(obesity_likelihood)

ggplot(tidy_obesity_likelihood, aes(x = term, y = estimate)) +
  geom_point(aes(color = estimate > 0), size = 4) +  # color points based on positive or negative coefficient
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.2) +  #error bars
  coord_flip() +  
  scale_color_manual(values = c("blue", "red")) +  # Set colors for positive and negative coefficients
  labs(title = "Obesity Likelihood",
       x = "Predictor Variables",
       y = "Effect on Log Odds of Being Considered Obese",
       color = "Effect Direction") +
  theme_minimal()  # Use a minimal theme

# Ensure the 'tidy_obesity_likelihood' contains the columns you want to summarize
tidy_obesity_likelihood <- broom::tidy(obesity_likelihood)

# Check available columns
names(tidy_obesity_likelihood)
```

The conclusions I can draw from @fig-coefficient align with what I expected given the dataset. It's clear that higher education levels, specifically 4-year degrees and post-grad degrees, are positively associated with support for Biden. Higher age groups, specifically 45-64 and 65+, are strongly tied to voting for Trump. Men are less likely to vote for Biden than Women.

\newpage
# Results

Figures @fig-page, @fig-pgender, and @fig-pedu are recreations of the earlier graphs we saw showcasing the proportion of voters for Joe Biden by groups within each demographic, where the bar is the data from CCES 2020 while the point is the prediction generated by the model.

```{r out.height="25%"}
#| label: fig-page
#| fig-cap: "Model prediction for votes by age vs. CCES 2020 data"
#| echo: false
#| warning: false
#| message: false
# Example constant values for the other factors
constant_education <- "2-year" # Choose a representative value
constant_gender <- "Male" # Choose a representative value

# Create a data frame for age predictions
age_data <- data.frame(age = c("18-29", "30-44", "45-64", "65+"),
                       education = constant_education,
                       gender = constant_gender)

# Predict probabilities for each age group
age_data$predicted_prob <- predict(obesity_likelihood, newdata = age_data, type = "response")
library(ggplot2)
# Actual proportions from the data
activity_data_grouped <- activity_data %>%
  group_by(age) %>%
  summarize(Proportion_for_Biden = mean(voted_for == "Biden", na.rm = TRUE))

# Assuming 'obesity_likelihood' is your logistic regression model object and
# 'age_data' contains the predicted probabilities with 'age' as a factor
age_data$age <- factor(age_data$age, levels = c("18-29", "30-44", "45-64", "65+"))

# Create a combined plot
ggplot() +
  geom_bar(data = activity_data_grouped, aes(x = age, y = Proportion_for_Biden, fill = age), 
           stat = "identity", position = "dodge", alpha = 0.5) +
  geom_point(data = age_data, aes(x = age, y = predicted_prob, color = age), 
             position = position_dodge(width = 0.9), size = 3) +
  scale_fill_brewer(palette = "Set1") + 
  scale_color_brewer(palette = "Set1") + 
  labs(title = "Actual vs. Predicted Support for Biden by Age Group",
       x = "Age Group", y = "Proportion") +
  theme_minimal() +
  theme(legend.position = "none")

```

```{r out.height="25%"}
#| label: fig-pgender
#| fig-cap: "Model prediction for votes by gender vs. CCES2020 data"
#| echo: false
#| warning: false
#| message: false
# Example constant values for the other factors (choose representative values for each)
constant_age <- "30-44" # Assuming this is a middle category for age

# Create a data frame for gender predictions
gender_data <- data.frame(
  gender = c("Male", "Female"),
  education = constant_education, # Assuming you have a constant value set for education
  age = constant_age
)

# Predict probabilities for each gender category
gender_data$predicted_prob <- predict(obesity_likelihood, newdata = gender_data, type = "response")

# Create a data frame for education predictions
education_data <- data.frame(
  education = c("No HS", "High school graduate", "Some college", "2-year", "4-year", "Post-grad"),
  gender = constant_gender, # Assuming you have a constant value set for gender
  age = constant_age
)

# Predict probabilities for each education category
education_data$predicted_prob <- predict(obesity_likelihood, newdata = education_data, type = "response")

# Actual proportions by gender
activity_data_gender_grouped <- activity_data %>%
  group_by(gender) %>%
  summarize(Proportion_for_Biden = mean(voted_for == "Biden", na.rm = TRUE))

# Actual proportions by education
activity_data_education_grouped <- activity_data %>%
  group_by(education) %>%
  summarize(Proportion_for_Biden = mean(voted_for == "Biden", na.rm = TRUE))
# Plot for Gender
ggplot() +
  geom_bar(data = activity_data_gender_grouped, aes(x = gender, y = Proportion_for_Biden, fill = gender), 
           stat = "identity", position = "dodge", alpha = 0.5) +
  geom_point(data = gender_data, aes(x = gender, y = predicted_prob, color = gender), 
             position = position_dodge(width = 0.9), size = 3) +
  scale_fill_brewer(palette = "Set2") + 
  scale_color_brewer(palette = "Set2") + 
  labs(title = "Actual vs. Predicted Support for Biden by Gender",
       x = "Gender", y = "Proportion") +
  theme_minimal() +
  theme(legend.position = "none")

```

```{r out.height="25%"}
#| label: fig-pedu
#| fig-cap: "Model prediction for votes by education level vs. CCES2020 data"
#| echo: false
#| warning: false
#| message: false
# Plot for Education
ggplot() +
  geom_bar(data = activity_data_education_grouped, aes(x = education, y = Proportion_for_Biden, fill = education), 
           stat = "identity", position = "dodge", alpha = 0.5) +
  geom_point(data = education_data, aes(x = education, y = predicted_prob, color = education), 
             position = position_dodge(width = 0.9), size = 3) +
  scale_fill_brewer(palette = "Set3") + 
  scale_color_brewer(palette = "Set3") + 
  labs(title = "Actual vs. Predicted Support for Biden by Education",
       x = "Education Level", y = "Proportion") +
  theme_minimal() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Adjust text angle for readability

```

By observation, it is evident that the model's prediction was fairly close to what the CCES data showed. The trend remains the same for each result. However, gender and education were predicted to have a higher sentiment of voting for Joe Biden while age had a lower sentiment.

It is worth nothing though that in @fig-page, all age categories showed polling data of above 50% to vote for Biden. Now with the lower average vote, the older age categories of "45-64" and "65+" are leaning towards Trump. On @fig-pedu for the lower education groups, the "No-HS" group which was at just about 50% is now above it, and the "High School Graduate" is also closer to voting for Biden. While this may seem marginal, close elections where voter sentiment is close to 50% would be largely impacted by changes like this. Although ultimately this is not an accurate predictor for who would win the election, only which groups would vote for who.

# Discussion

This analysis demonstrates the demographics play in political preferences, with results that align with the existing literature. Our data suggests that age, education, and gender significantly influence electoral outcomes. In particular, the model's predictions closely mirrored CCES 2020 data trends, reinforcing the importance of these variables in understanding voter behavior. High education and younger ages showed high correlation with voting for joe biden.

It is important to note that based on the sample size of our data (as seen in @fig-education and @fig-agegender), the majority age groups which participated in the survey were "45-64" and "65+", while the variables of education and gender were much more closer together. If this were representative of the actual population then it would suggest that the age groups are skewed more towards those that support Trump, so despite most of the proportions being towards Biden, it can be a lot more nuanced of there are more groups in demographics that support Trump.

The demographic trends observed in our analysis, specifically the tendency of younger voters, women, and individuals with higher educational attainment towards supporting Joe Biden, mirror broader societal shifts. These findings align with historical patterns where demographic attributes have increasingly become predictors of political alignment. For instance, the growing political polarization along educational lines has been well-documented, with individuals possessing higher educational qualifications tending to support more progressive policies and candidates [@Dalton2020].

Moreover, the gender gap in political preferences is not a novel phenomenon. Women have been consistently more likely to vote for Democratic candidates in recent elections, a trend that has been attributed to differing policy priorities, including healthcare, reproductive rights, and social welfare [@Schaffner2018]. The 2020 election further highlighted this divide, as issues such as women's rights and gender equality took center stage, prompting a significant mobilization among female voters [@Coffe2020].

Age demographics have also played a critical role in shaping the political landscape. The engagement and political leanings of younger voters have been subjects of extensive study, particularly given their inclination towards progressive and liberal policies. This has been partly driven by generational differences in value systems, with younger cohorts exhibiting greater concern for issues such as climate change, social justice, and economic inequality [@Norris2021]. The 2020 election underscored the importance of the youth vote, which showed overwhelming support for Biden, reflecting a broader trend of political mobilization among younger Americans in response to current socio-political challenges [@Smith2020].

The implications of these demographic trends extend beyond electoral outcomes. They signal shifting societal values and priorities, with potential long-term impacts on policy direction and political discourse. For instance, the increasing support for progressive policies among younger and more educated demographics may influence the policy agendas of political parties and candidates in future elections [@Inglehart2020].

It's also crucial to note the limitations of relying solely on demographic variables to predict electoral outcomes. While age, education, and gender provide significant insights, the complex interplay of socio-economic factors, geographic location, and individual experiences also plays a crucial role in shaping political preferences. Future research should aim to incorporate a broader range of variables to capture the complexity of voter behavior more accurately.

Our analysis contributes to the growing body of literature on demographic influences on political preferences, offering valuable insights into the dynamics of the 2020 U.S. presidential election. As the nation continues to navigate through periods of significant social and political change, understanding the demographic underpinnings of electoral outcomes will remain a critical area of study.

# Weaknesses and Next Steps
Although valuable, this study provides an intriguing analysis of how demographic influences affect political preferences; however, there are additional limitations to consider. Firstly, the study's analysis model, which used logistic regression, may have simplified the understanding of complex dynamics and could have overlooked some interactions between demographic variables that would provide a better understanding of voter behavior. Future research could utilize more advanced tools, such as multilevel or mixed-effects models, to more accurately capture the interplay of these demographic variables.

Secondly, while detailed, the study's dataset may not fully account for the subtleties in voter behavior across all demographics. For example, the impact of regional urban-rural political contexts on voting patterns could have nuances beyond the current analysis's reach. Incorporating geographic and socioeconomic variables could further improve and strengthen the model, offering deeper insight into the electoral impact of demographic factors.

Lastly, the study's focus on the 2020 election, though timely, limits its ability to capture longitudinal trends in voter behavior. A comparative analysis developed over multiple election cycles might reveal changes in demographic influences on political preferences over time, providing a richer context for interpreting the 2020 election results. In summary, this paper not only sheds light on the demographic determinants of voter preferences in a pivotal election year but also encourages reflection on the broader societal shifts these patterns indicate. As America faces deep social and political challenges, understanding the demographic underpinnings of political allegiance becomes increasingly important. Further research, building on the insights provided in this paper and addressing the weaknesses outlined, has the potential to offer even better explanations of the complex tapestry of American electoral politics.
\newpage

# References